---
title: "Psychometric Models in `irt` Package"
author: "Emre Gönülateş"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
vignette: >
  %\VignetteIndexEntry{Psychometric Models in `irt` Package}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  collapse = TRUE,
  comment = "#>"
)
```

This vignettes covers the psychometric models that are implemented in 
the `irt` package. 


# Item Models

| Name  | Description                                     | Parameters    |
|-------|-------------------------------------------------|---------------|
| Rasch | Rasch Model                                     | b             |
| 1PL   | One-Parameter Logistic Model                    | b, D          |
| 2PL   | Two-Parameter Logistic Model                    | a, b, D       |
| 3PL   | Three-Parameter Logistic Model                  | a, b, c, D    |
| 4PL   | Four-Parameter Logistic Model                   | a, b, c, d, D |
| GRM   | Graded Response Model                           | a, b, D       |
| PCM   | Partial Credit Model                            | b             |
| GPCM  | Generalized Partial Credit Model                | a, b, D       |
| GPCM2 | Reparametrized Generalized Partial Credit Model | a, b, d, D    |


## Rasch Model

For an examinee $i$ with ability $\theta_i$, the probability of correct 
response to an item $j$ is: 

$$P\left(X_{ij} = 1 | \theta_i;b_j\right) = 
  \frac{e^{(\theta_i - b_j)}}{1 + e^{(\theta_i - b_j)}}
$$

where , $b_j$ is the item difficulty (or threshold) of item $j$.  

User needs to specify only the item difficulty parameter: 

```{r}
library(irt)

itm_rasch <- item(b = -1.29)
itm_rasch
```

The probability of each response option at $\theta = -.65$ is: 

```{r}
prob(ip = itm_rasch, theta = -0.65)
```

The item characteristic curve of this item is: 
```{r}
plot(itm_rasch)
```

<!-- ####################################################################### -->

## One-Parameter Logistic Model

For an examinee $i$ with ability $\theta_i$, the probability of correct 
response to an item $j$ is: 

$$P\left(X_{ij} = 1 | \theta_i;b_j\right) = 
  \frac{e^{D(\theta_i - b_j)}}{1 + e^{D(\theta_i - b_j)}}
$$

where , $b_j$ is the item difficulty (or threshold) of item $j$. D is the
scaling constant (the default value is `r irt:::DEFAULT_D_SCALING`). 

User needs to specify the following parameters: 

```{r}
itm_1pl <- item(b = 0.83, D = 1)
itm_1pl
```

The probability of each response option at $\theta = .73$ is: 

```{r}
prob(ip = itm_1pl, theta = 0.73)
```

The item characteristic curve of this item is: 
```{r}
plot(itm_1pl)
```

<!-- ####################################################################### -->

## Two-Parameter Logistic Model

For an examinee $i$ with ability $\theta_i$, the probability of correct 
response to an item $j$ is: 

$$P\left(X_{ij} = 1 | \theta_i;a_j, b_j\right) = 
  \frac{e^{Da_j(\theta_i - b_j)}}{1 + e^{Da_j(\theta_i - b_j)}}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_j$ is the item 
difficulty (or threshold). D is the scaling constant (the default value is 
`r irt:::DEFAULT_D_SCALING`). 

User needs to specify all of the parameter values: 

```{r}
itm_2pl <- item(a = .94, b = -1.302, D = 1)
itm_2pl
```

The probability of each response option at $\theta = -0.53$ is: 

```{r}
prob(ip = itm_2pl, theta = -0.53)
```

The item characteristic curve of this item is: 
```{r}
plot(itm_2pl)
```

<!-- ####################################################################### -->

## Three-Parameter Logistic Model

For an examinee $i$ with ability $\theta_i$, the probability of correct 
response to an item $j$ is: 

$$P\left(X_{ij} = 1 | \theta_i;a_j, b_j, c_j\right) = 
  c_j + (1 - c_j)\frac{e^{Da_j(\theta_i - b_j)}}{1 + e^{Da_j(\theta_i - b_j)}}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_j$ is the item 
difficulty (or threshold), and $c_j$ is the lower asymptote (or pseudo-guessing) 
value. D is the scaling constant (the default value is 
`r irt:::DEFAULT_D_SCALING`). 

As can be seen from the equation above, the user needs to specify all of the 
parameter values: 

```{r}
itm_3pl <- item(a = 1.51, b = 2.04, c = .16, D = 1.7)
itm_3pl
```

The probability of each response option at $\theta = 1.5$ is: 

```{r}
prob(ip = itm_3pl, theta = 1.5)
```

The item characteristic curve of this item: 
```{r}
plot(itm_3pl)
```

<!-- ####################################################################### -->

## Four-Parameter Logistic Model

For an examinee $i$ with ability $\theta_i$, the probability of correct 
response to an item $j$ is: 

$$P\left(X_{ij} = 1 | \theta_i;a_j, b_j, c_j, d_j\right) = 
  c_j + (d_j - c_j)\frac{e^{Da_j(\theta_i - b_j)}}{1 + e^{Da_j(\theta_i - b_j)}}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_j$ is the item
difficulty (or threshold), $c_j$ is the lower asymptote (or pseudo-guessing)
value and $d_j$ is the upper asymptote. D is the scaling constant (the default
value is `r irt:::DEFAULT_D_SCALING`). 
 
As can be seen from the equation above, the user needs to specify all of the 
parameter values: 

```{r}
itm_4pl <- item(a = 1.2, b = -.74, c = .22, d = .99, D = 1.7)
itm_4pl
```

The probability of each response option at $\theta = 1.2$ is: 

```{r}
prob(ip = itm_4pl, theta = 1.2)
```

The item characteristic curve of this item is: 
```{r}
plot(itm_4pl)
```

<!-- ####################################################################### -->

## Graded Response Model (GRM)

For an examinee $i$ with ability $\theta_i$, the probability of responding at 
or above the category $k$ to an item $j$ with possible scores 
$k = 0, 1, \ldots, m_j$: 

$$P^*\left(X_{ij} = k | \theta_i;a_j, b_j\right) =
  \frac{e^{Da_j(\theta_i - b_{jk})} }{1 + e^{Da_j(\theta_i - b_{jk})}}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_{jk}$ is the 
threshold parameter. Note that the probability of responding at or above the
lowest category is $P^*\left(X_{ij} = 0\right) = 1$. Responding at a category 
$k$ can be calculated as:

$$P(X_{ij}=k|\theta_i) = P^*(X_{ij} = k) - P^*(X_{ij} = k+1)$$
 
The user needs to specify the following parameter values: 

```{r}
itm_grm <- item(a = 0.84, b = c(-1, -.2, .75, 1.78), D = 1.7, model = "GRM")
itm_grm
```

The probability of each response option at $\theta = 1.13$ is: 

```{r}
prob(ip = itm_grm, theta = 1.13)
```

The option characteristic curves of this item is: 
```{r}
plot(itm_grm)
```


<!-- ####################################################################### -->

## Generalized Partial Credit Model (GPCM)

For an examinee $i$ with ability $\theta_i$, the probability of a 
response $k$ to an item $j$ with possible scores $k = 0, 1, \ldots, m_j$: 

$$P\left(X_{ij} = k | \theta_i;a_j, b_j\right) =
  \frac{\text{exp} \left( \sum_{v = 0}^{k}Da_j(\theta_i - b_{jv}) \right) }
  {{\sum_{h = 0}^{m_j} \text{exp} \left[ \sum_{v = 0}^{h}
     Da_j(\theta_i - b_{jv})  \right] }}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_{jv}$ are the 
step difficulty parameters. Note that $b_{jv}$ values are not necessarily 
ordered from smallest to the largest. D is the scaling constant (the default
value is `r irt:::DEFAULT_D_SCALING`). 
$\sum_{v = 0}^0 Da_j(\theta_i - b_{jv}) = 0$. 
 
The user needs to specify the following parameter values: 

```{r}
itm_gpcm <- item(a = 1.1, b = c(-.74, .3, .91, 2.19), D = 1.7, model = "GPCM")
itm_gpcm
```

The probability of each response option at $\theta = -0.53$ is: 

```{r}
prob(ip = itm_gpcm, theta = -0.53)
```

The option characteristic curves of this item is: 
```{r}
plot(itm_gpcm)
```

<!-- ####################################################################### -->

## Partial Credit Model (PCM)

For an examinee $i$ with ability $\theta_i$, the probability of a  
response $k$ to an item $j$ with possible scores $k = 0, 1, \ldots, m_j$: 

$$P\left(X_{ij} = k | \theta_i;b_j\right) =
  \frac{\text{exp} \left( \sum_{v = 0}^{k}(\theta_i - b_{jv}) \right) }
  {{\sum_{h = 0}^{m_j} \text{exp} \left[ \sum_{v = 0}^{h}
     (\theta_i - b_{jv})  \right] }}
$$

where $b_{jv}$ are the step difficulty parameters. 
$\sum_{v = 0}^0 (\theta_i - b_{jv}) = 0$. 
 
The user needs to specify the following parameter values: 

```{r}
itm_pcm <- item(b = c(-1.38, -.18, 1.1), model = "PCM")
itm_pcm
```

The probability of each response option at $\theta = -1.09$ is: 

```{r}
prob(ip = itm_pcm, theta = -1.09)
```

The option characteristic curves of this item is: 
```{r}
plot(itm_pcm)
```



<!-- ####################################################################### -->

## Re-parametrized Generalized Partial Credit Model (GPCM2)

For an examinee $i$ with ability $\theta_i$, the probability of a 
response $k$ to an item $j$ with possible scores $k = 0, 1, \ldots, m_j$: 

$$P\left(X_{ij} = k | \theta_i;a_j, b_j, d_j\right) =
  \frac{\text{exp} \left( \sum_{v = 0}^{k}Da_j(\theta_i - b_j + d_{jv}) \right) }
  {{\sum_{h = 0}^{m_j} \text{exp} \left[ \sum_{v = 0}^{h}
     Da_j(\theta_i - b_j + d_{jv})  \right] }}
$$

where $a_j$ is the item discrimination (or slope) of item $j$, $b_j$ is the 
overall location parameter and $d_{jv}$ are the threshold parameters. 
D is the scaling constant (the default value is `r irt:::DEFAULT_D_SCALING`). 
 
The user needs to specify the following parameter values: 

```{r}
itm_gpcm2 <- item(a = .71, b = .37, d = c(-.18, .11, 1.29), D = 1, 
                  model = "GPCM2")
itm_gpcm2
```

The probability of each response option at $\theta  1.3$ is: 

```{r}
prob(ip = itm_gpcm2, theta = 1.3)
```

The option characteristic curves of this item is: 
```{r}
plot(itm_gpcm2)
```
